Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-10 main.tex
Make some modifications in the graph part

\newcommand{\graph}[1]{\mathcal{#1}}
\newcommand{\vertices}[1][]{\ifthenelse{\isempty{#1}}{V}{V_{\graph{#1}}}}
\newcommand{\edges}[1][]{\ifthenelse{\isempty{#1}}{E}{E_{\graph{#1}}}}
\newcommand{\bgraph}[1]{\mathcal{\vec{#1}}}

@@

@@

\newcommand{\Gr}{\mathcal{G}}
\newcommand{\LG}{\mathcal{L}}
\newcommand{\FT}{\mathcal{F}}
\newcommand{\US}{\mathcal{U}}
\newcommand{\PE}[1]{#1^P}

@@

Let $\gG=\tuple{\vG,\eG}$ and $\gH=\tuple{\vH,\eH}$ be graphs such
that $\vG\cap\vH=\emptyset$. A \bfit{homomorphism}
$\phi\colon\gG\to\gH$ is a function $\phi\colon\vG\to\vH$ such that if
$vw\in\eG$ then $\phi(v)\phi(w)\in\eH$. The \bfit{union} $\gG +\gH$ is
the graph $\tuple{\vG \cup\vH,\eG\cup \eH}$ and the \bfit{join} $\gG
\times \gH$ is the graph $\tuple{\vG \cup\vH,\eG \cup \eH \cup \set{vw
    | v \in \vG, w \in \vH}}$.  A graph $\gG$ is \bfit{disconnected}
if $\gG=\gG_1+\gG_2$ for two non-empty graphs $\gG_1,\gG_2$, otherwise
it is \bfit{connected}. It is \emph{coconnected} if its complement is
connected.

@@

\begin{thm}[\cite{}]\label{thm:cograph}
  A graph is a cograph iff it can be constructed from the singletons via the operations $+$ and~$\times$.
\end{thm}
\todo{ad the reference}

In a graph $\gG$, the \bfit{neighbourhood} $N(v)$ of a vertex $v \in\vG$ is defined as
the set $\set{w | vw \in\eG}$. 
A \bfit{module} is a set $M \subseteq \vG$ such that $N(v) \ M = N(w) \ M$
for all $v, w \in M$. A module $M$ is \bfit{strong} if for every
module $M'$, we have $M' \subseteq M$, $M \subseteq$ or $M \cap M' =
\emptyset$. A module is \bfit{proper} if it has two or more vertices.

@@

A cograph is \bfit{logical} if every vertex is labelled either by an
atom or variable, and it has at least one atom-labelled vertex. We
write $\single\alpha$ for an $\alpha$-labelled vertex. An
atom-labelled vertex is called a \bfit{literal} and a
variable-labelled vertex is called a \bfit{binder}. A binder labelled
with $x$ is called an \bfit{$x$-binder}. The \bfit{scope} of a binder
$b$ is the smallest proper strong module containing $b$.
An \bfit{$x$-literal} is a literal whose
atom contains the variable $x$. An $x$-binder \bfit{binds} every
$x$-literal in its scope.  In a logical cograph $G$, a binder $b$ is
\bfit{existential} (resp. \bfit{universal}) if, for every other vertex
$v$ in its scope, we have $bv \in E_G$ (resp. $bv \notin E_G$). An
$x$-binder is \bfit{legal} if its scope contains no other $x$-binder
and at least one literal.

@@

For a formula $A$, we can define its
associated fograph $\graphof(A)$ inductively by: $\graphof(\alpha) =
\cdot \alpha$ (a single-vertex graph whose vertex is labelled by
$\alpha$); $\graphof(A \cor B) = \graphof(A) + \graphof(B)$ and
$\graphof(A \cand B) = \graphof(A) \times \graphof(B)$.
\todo{make look nice, use array}

@@

@@

Two atoms are \bfit{pre-dual} if their predicate symbols are dual
(e.g. $pxy$ and $\overline{p}yz$) and two literals are \bfit{pre-dual} if their
labels (atoms) are pre-dual. A \bfit{linked fograph} is a coloured fograph such
that every colour, called a \bfit{link}, consists of two pre-dual literals, and
every literal is either $\ttt$-labelled or in a link.

Let $G$ be a linked fograph. The set of links can be seen as a unification problem
by identifying dual predicate symbols. A \bfit{dualizer} of $G$ is an assignment
unifying all the links of $G$. Since a first-order unification problem is either
solvable or has a most general unifier, we can define the notion of \bfit{most
general dualizer}. A \bfit{dependency} is a pair $(\cdot x, \cdot y)$ of an
existential binder $\cdot x$ and a universal binder $\cdot y$ such that the most
general dualizer assigns to $x$ a term containing $y$. A \bfit{leap} is either a
link or a dependency. The \bfit{leap graph} $\LG_G$ of $G$ is the undirected graph $(V_G,
L_G)$ where $L_G$ is the set of leaps of $G$. A set $W \subseteq V_G$ induces a
\bfit{bimatching} in $G$ if $W$ induces a matching in $G$ and in the leap graph
$\LG_G$ as well.

\todo{define matching}

>>>>

\newcommand{\graph}[1]{\mathcal{#1}}
\newcommand{\vertices}[1][]{\ifthenelse{\isempty{#1}}{V}{V_{\graph{#1}}}}
\newcommand{\edges}[1][]{\ifthenelse{\isempty{#1}}{E}{E_{\graph{#1}}}}
\newcommand{\bgraph}[1]{\mathcal{\vec{#1}}}
\newcommand{\lgraph}[1]{\mathcal{L_{#1}}}
\newcommand{\leaps}[1]{L_{#1}}

@@

\newcommand{\lgG}{\lgraph{\gG}}

\newcommand{\lpG}{\leaps{\gG}}

@@

\newcommand{\Gr}{\mathcal{G}}
\newcommand{\FT}{\mathcal{F}}
\newcommand{\US}{\mathcal{U}}
\newcommand{\PE}[1]{#1^P}

@@

Let $\gG=\tuple{\vG,\eG}$ and $\gH=\tuple{\vH,\eH}$ be graphs such
that $\vG\cap\vH=\emptyset$. A \bfit{homomorphism}
$\phi\colon\gG\to\gH$ is a function $\phi\colon\vG\to\vH$ such that if
$vw\in\eG$ then $\phi(v)\phi(w)\in\eH$. The \bfit{union} $\gG +\gH$ is
the graph $\tuple{\vG \cup\vH,\eG\cup \eH}$ and the \bfit{join} $\gG
\times \gH$ is the graph $\tuple{\vG \cup\vH,\eG \cup \eH \cup \set{vw
\mid v \in \vG, w \in \vH}}$.  A graph $\gG$ is \bfit{disconnected}
if $\gG=\gG_1+\gG_2$ for two non-empty graphs $\gG_1,\gG_2$, otherwise
it is \bfit{connected}. It is \emph{coconnected} if its complement is
connected.

@@

\begin{thm}[\cite{duffin:65}]\label{thm:cograph}
  A graph is a cograph iff it can be constructed from the singletons via the operations $+$ and~$\times$.
\end{thm}
\todo{ad the reference} \juihsuan{done}

In a graph $\gG$, the \bfit{neighbourhood} $N(v)$ of a vertex $v \in\vG$ is defined as
the set $\set{w \mid vw \in\eG}$. A set $W \subseteq \vG$ induces a \bfit{matching} in
$\gG$ if for all $w \in W$, $N(w) \cap W$ is a singleton.
A \bfit{module} is a set $M \subseteq \vG$ such that $N(v) \ M = N(w) \ M$
for all $v, w \in M$. A module $M$ is \bfit{strong} if for every
module $M'$, we have $M' \subseteq M$, $M \subseteq$ or $M \cap M' =
\emptyset$. A module is \bfit{proper} if it has two or more vertices.

@@

A cograph is \bfit{logical} if every vertex is labelled either by an
atom or variable, and it has at least one atom-labelled vertex. We
write $\single\alpha$ for an $\alpha$-labelled vertex. An
atom-labelled vertex is called a \bfit{literal} and a
variable-labelled vertex is called a \bfit{binder}. A binder labelled
with $x$ is called an \bfit{$x$-binder}. The \bfit{scope} of a binder
$b$ is the smallest proper strong module containing $b$.
An \bfit{$x$-literal} is a literal whose
atom contains the variable $x$. An $x$-binder \bfit{binds} every
$x$-literal in its scope.  In a logical cograph $\gG$, a binder $b$ is
\bfit{existential} (resp. \bfit{universal}) if, for every other vertex
$v$ in its scope, we have $bv \in \eG$ (resp. $bv \notin \eG$). An
$x$-binder is \bfit{legal} if its scope contains no other $x$-binder
and at least one literal.

@@

For a formula $A$, we can define its
associated fograph $\graphof(A)$ inductively by: 
\begin{equation*}
  \begin{array}{l}
    \graphof{\alpha} = \cdot \alpha (\text{single-vertex graph whose vertex is 
    labelled by } \alpha); \\[.5ex]
    \graphof{A \cor B} = \graphof{A} + \graphof{B};
    \graphof{A \cand B} = \graphof{A} \times \graphof{B}; \\[.5ex]
    \graphof{\forall x . A} = \cdot x + \graphof{A};
    \graphof{\exists x . A} = \cdot x \times \graphof{A}.
  \end{array}
\end{equation*}
\todo{make look nice, use array} \juihsuan{done}

@@

We define a congruence relation on formulas by the following equations:
\begin{equation*}
  \begin{array}{cc}
    A \vlan B \fequ B \vlan A & \forall x . (A \vlor B) \fequ (\forall x . A)
    \vlor B \\
    A \vlor B \fequ B \vlor A & \exists x . (A \vlan B) \fequ (\exists x . A)
    \vlan B \\
    (A \vlan B) \vlan C \fequ A \vlan (B \vlan C) & \forall x . \forall y . A
    \fequ \forall y . \forall x . A \\
    (A \vlor B) \vlor C \fequ A \vlor (B \vlor C) & \exists x . \exists y . A
    \fequ \exists y . \exists x . A \\
  \end{array}
\end{equation*}

@@

Two atoms are \bfit{pre-dual} if their predicate symbols are dual
(e.g. $p(x, y)$ and $\dual{p}(y, z)$) and two literals are \bfit{pre-dual} if their
labels (atoms) are pre-dual. A \bfit{linked fograph} is a coloured fograph such
that every colour, called a \bfit{link}, consists of two pre-dual literals, and
every literal is either $\ttt$-labelled or in a link.

Let $\gG$ be a linked fograph. The set of links can be seen as a unification problem
by identifying dual predicate symbols. A \bfit{dualizer} of $\gG$ is an assignment
unifying all the links of $\gG$. Since a first-order unification problem is either
solvable or has a most general unifier, we can define the notion of \bfit{most
general dualizer}. A \bfit{dependency} is a pair $(\cdot x, \cdot y)$ of an
existential binder $\cdot x$ and a universal binder $\cdot y$ such that the most
general dualizer assigns to $x$ a term containing $y$. A \bfit{leap} is either a
link or a dependency. The \bfit{leap graph} $\lgG$ of $\gG$ is the undirected
graph $(\vG, \lpG)$ where $\lpG$ is the set of leaps of $\gG$. A set $W
\subseteq \vG$ induces a
\bfit{bimatching} in $\gG$ if $W$ induces a matching in $\gG$ and in the leap graph
$\lgG$ as well.

\todo{define matching} \juihsuan{done in the cograph part}
